---
title: "Take Home Exercise 2"
description: |
  In this Take-home Exericse, (A) an investigation will be conducted to see if the distribution of Airbnb listings is affected by location factors and (B) analyse the impact of COVID-19 on Airbnb business in Singapore by comparing Airbnb listings data on June 2019 and June 2021.
author:
  - name: Ngah Xin Yan
    url: https://github.com/nxinyan/
date: 09-26-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina = 3)
```

## Overview

The 'sharing economy' has seen unparalleled growth in terms of number of users, making it possible for new methods of economic and social interactions. 

Airbnb, once belonging to the sharing economy, has expanded into many cities across many countries and successfully transitioned into profitability, showing the valid validity of sharing economy within the global market.

Singapore has yet to legalise short-term rentals offered by platforms such as Airbnb. However, there are datasets for Airbnb listing in Singapore.

## Objective
### Section A: Airbnb Distribution

To investigate if the distribution of Airbnb listings are affected by location factors 

### Section B: Impact of COVID-19

To analyse the impact of COVID-19 on Airbnb business in Singapore by comparing Airbnb listings data on June 2019 and June 2021.

## Data Used

- **Airbnb Listing (June 2019)**: Singapore's Airbnb listing for 2019 (csv file) from [Inside Airbnb](http://insideairbnb.com/get-the-data.html)
- **Airbnb Listing (July 2021)**: Singapore's Airbnb listing for 2021 (csv file) from [Inside Airbnb](http://insideairbnb.com/get-the-data.html)
- **Hotels**: Listing of hotels in Singapore (csv file) from [OneMap API](https://cran.r-project.org/web/packages/onemapsgapi/index.html)
- **Tourism**: List of tourist attraction in Singapore (csv file) from [OneMap API](https://cran.r-project.org/web/packages/onemapsgapi/index.html)
- **CostalOutline**: National Boundary of Singapore (shp file) from [data.gov.sg](https://data.gov.sg/dataset/national-map-polygon)
- **MP14_SUBZONE_WEB_PL**: URA 2014 Master Plan Planning Subzone Boundary Data (shp file) from [data.gov.sg](https://data.gov.sg/dataset/master-plan-2014-subzone-boundary-web)
- **MRTLRTStnPtt**: All MRT and LRT stations in Singapore (shp file) from [LTA Datamall](https://datamall.lta.gov.sg/content/datamall/en/search_datasets.html?searchText=mrt)


## Installing and Loading the R packages

- **sf**: used to import, manage and process vector-based geospatial data in R.

- [**spatstat**](https://spatstat.org/): for point pattern analysis.

- [**raster**](https://cran.r-project.org/web/packages/raster/): reads, writes, manipulates, analyses and model of gridded spatial data.

- [**maptools**](https://cran.r-project.org/web/packages/maptools/index.html): provides a set of tools for manipulating geographic data.

- [**tmap**](https://cran.r-project.org/web/packages/tmap/index.html): to plot cartographic quality static point patterns maps or interactive maps by using leaflet API.

- **tidyverse**: to perform data science tasks such as importing, wrangling and visualising data

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap','tidyverse','rgdal','ggplot2', 'ggthemes', 'plotly')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Data Extraction, Wrangling and Integration
### Importing Geospatial Data

Importing the following geospatial dataset into R by using *st_read()* of **sf** package

**`MP14_SUBZONE_WEB_PL`**

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

**`CostalOutline`**

```{r}
sg <- st_read(dsn = "data/geospatial", layer="CostalOutline")
```

**`MRTLRTStnPtt`**

```{r}
mrtlrt <- st_read(dsn = "data/geospatial", layer="MRTLRTStnPtt")
```
The projected CRS of all the data frames are SVY21, which is relevant for analysis in Singapore. However, the EPSG code should be checked as when importing geospatial data into R, the coordinate system of the source data can go missing or be wrongly assigned.

### Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
```

```{r}
st_crs(sg)
```

```{r}
st_crs(mrtlrt)
```

All data frames indicated that the EPSG code is 9001 which is wrong. The EPSG code for SVY21 is 3414.

*st_crs(sg)* of **sf** package will be used to assign the correct EPSG code to all data frames.

```{r}
mpsz_sf <- st_transform(mpsz, 3414)
```

```{r}
sg_sf <- st_transform(sg, 3414)
```

```{r}
mrtlrt_sf <- st_transform(mrtlrt, 3414)
```

Checking CRS again

```{r}
st_crs(mpsz_sf)
```

```{r}
st_crs(sg_sf)
```

```{r}
st_crs(mrtlrt_sf)
```

All dataframes has the EPSG code of 3414 now.

### Importing Aspatial Data 

Importing the following aspatial dataset into R dataframe by using *read_csv()* function of **readr** package in **tidyverse**:

```{r}
airbnb_june2019 <- read_csv("data/aspatial/airbnb_june2019.csv")
airbnb_july2021 <- read_csv("data/aspatial/airbnb_july2021.csv")
hotels <- read_csv("data/aspatial/hotels.csv")
tourism <- read_csv("data/aspatial/tourism.csv")
```

Use ***glimpse()*** to learn more about the attribute information in the data frame.

```{r}
glimpse(airbnb_june2019)
```

```{r}
glimpse(airbnb_july2021)
```

```{r}
glimpse(hotels)
```

```{r}
glimpse(tourism)
```

***glimpse()*** report reveals the data type of each fields. 

The ***latitude*** and ***longitude*** column are for the coordinates, which are in decimal degree format. There are also some dataset which uses ***LATITUDE***/***LONGITUDE*** and ***longitude***/***Lng*** columns instead, but it still shows the coordinates in decimal degree format. It is good to assume that the data is in **WGS84** . It has to be assigned to the correct coordinate system to match the geospatial data above.  

### Pre-Processing of Data 
**Dealing with Missing Values**

Using the ***is.na()*** function, check for missing value in the ***latitude***/***Lat*** and ***longitude***/***Lng*** column.

```{r}
any(is.na(airbnb_june2019$latitude))
any(is.na(airbnb_june2019$longitude))
```

```{r}
any(is.na(airbnb_july2021$latitude))
any(is.na(airbnb_july2021$longitude))
```

```{r}
any(is.na(hotels$Lat))
any(is.na(hotels$Lng))
```

```{r}
any(is.na(tourism$LATITUDE))
any(is.na(tourism$LONGTITUDE))
```

```{r}
any(is.na(tourism$Lat))
any(is.na(tourism$Lng))
```

There are no NA values for ***airbnb_june2019***, ***airbnb_july2021 ***, ***hotels *** dataset for Latitude and Longitude. However, for  ***tourism ***, there is missing value in its  ***LATITUDE*** and  ***Longtitude*** column. However, there is no missing value in its ***Lat*** and ***Lng*** column. Thus, the ***Lat*** and ***Lng*** column will be used to create the simple data frame instead.

```{r}
tourism <- tourism %>%
            drop_na("LATITUDE")

tourism <- tourism %>%
            drop_na("LONGTITUDE")
```

### Transforming the projection of coordinate system

All the data will be converted into simple feature data frame using ***st_as_sf()*** function. For the coords argument, ***longitude*** will be used as the x-coordinate and ***latitude*** will be the y-coordinate. WGS84 geographic coordinate system uses EPSG code 4326 but we have to match our geospatial data projected coordinate system. %<% is used to nest ***st_transform()*** trabsform the new simple feature data frame in SVY21 (EPSG: 3414) projected coordinates system. 

```{r}
airbnb_june2019_sf <- st_as_sf(airbnb_june2019, 
                          coords = c("longitude", 
                                     "latitude"), 
                          crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
airbnb_july2021_sf <- st_as_sf(airbnb_july2021, 
                          coords = c("longitude", 
                                     "latitude"), 
                          crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
hotels_sf <- st_as_sf(hotels, 
                          coords = c("Lng", 
                                     "Lat"), 
                          crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
tourism_sf <- st_as_sf(tourism, 
                          coords = c("LONGTITUDE", 
                                     "LATITUDE"), 
                          crs = 4326) %>%
  st_transform(crs=3414)
```

Checking CRS again

```{r}
st_crs(airbnb_june2019_sf)
st_crs(airbnb_july2021_sf)
st_crs(hotels_sf)
st_crs(tourism_sf)
```

All the data has successfully been changed to EPSG 3414.

## Geospatial Data Wrangling

Both geospatial and aspatial data will be converted from simple feature frame to sp's Spatial* class.

### Converting sf data frames to sp’s Spatial* class

*as_Spatial()* is used to convert the all the geospatial data from simple feature data frame to sp’s Spatial* class.

```{r}
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
mrtlrt <-as_Spatial(mrtlrt_sf)
                    
airbnb_june2019 <- as_Spatial(airbnb_june2019_sf)
airbnb_july2021 <- as_Spatial(airbnb_july2021_sf)
hotels <- as_Spatial(hotels_sf)
tourism <- as_Spatial(tourism_sf)                    
```

Checking all the Spatial* class

```{r}
mpsz
```

```{r}
sg
```

```{r}
mrtlrt
```

```{r}
airbnb_june2019
```

```{r}
airbnb_july2021
```

```{r}
hotels
```

```{r}
tourism
```

All the geospatial data have been converted to their respective sp's Spatial* classes.

### Converting the Spatial* class into generic sp format

It is mandatory for the analytical data to be in ***ppp*** object form to use **spatstat**. Since there is no direct way to convert a Spatial* classes into ***ppp*** object, ***Spatial classes**** needs to be converted into ***Spatial*** object first.

```{r}
mpsz_sp <- as(mpsz, "SpatialPolygons")
sg_sp <- as(sg, "SpatialPolygons")
mrtlrt_sp <- as(mrtlrt, "SpatialPoints")

airbnb_june2019_sp <- as(airbnb_june2019, "SpatialPoints")
airbnb_july2021_sp <- as(airbnb_july2021, "SpatialPoints")
hotels_sp <- as(hotels, "SpatialPoints")
tourism_sp <- as(tourism, "SpatialPoints")
```

### Converting the generic sp format into spatstat’s ppp format

*as.ppp()* function of **spatspat** is used to convert the spatial data into **spatstat**'s ***ppp*** object format.

```{r}
mrtlrt_ppp <- as(mrtlrt_sp, "ppp")
airbnb_june2019_ppp <- as(airbnb_june2019_sp, "ppp")
airbnb_july2021_ppp <- as(airbnb_july2021_sp, "ppp")
hotels_ppp <- as(hotels_sp, "ppp")
tourism_ppp <- as(tourism_sp, "ppp")
```

Taking a quick look at the summary statistics of the newly created **ppp** object with *summary()* 

```{r}
summary(mrtlrt_ppp)
```

```{r}
summary(airbnb_june2019_ppp)
```

```{r}
summary(airbnb_july2021_ppp)
```

```{r}
summary(hotels_ppp)
```

```{r}
summary(tourism_ppp)
```

From the summary above, there are 3 ppp objects that contains duplicated points: airbnb_june2019_ppp, airbnb_july2021_ppp, hotels_ppp and tourism_ppp

### Handling Duplicated Points

Checking for duplication in ***ppp*** object and the number of duplicated points

```{r}
any(duplicated(mrtlrt_ppp))
sum(multiplicity(mrtlrt_ppp) > 1)
```

There are no duplicated points in mrtlrt_ppp. 

```{r}
any(duplicated(airbnb_june2019_ppp))
sum(multiplicity(airbnb_june2019_ppp) > 1)
```

There are 6 duplicated points in *airbnb_june2019_ppp* as shown above.

```{r}
any(duplicated(airbnb_july2021_ppp))
sum(multiplicity(airbnb_july2021_ppp) > 1)
```

There are 224 duplicated points in *airbnb_july2021_ppp* as shown above.

```{r}
any(duplicated(hotels_ppp))
sum(multiplicity(hotels_ppp) > 1)
```

There are 10 duplicated points in *hotels_ppp* as shown above.

```{r}
any(duplicated(tourism_ppp))
sum(multiplicity(tourism_ppp) > 1)
```

There are 7 duplicated points in *tourism_ppp* as shown above.

### Jittering

To overcome the problem of duplicates, *jittering* will be used. Jittering adds a small perturbation to the duplicate points so that they do not occupy the exact same space.

```{r}
airbnb_june2019_ppp_jit <- rjitter(airbnb_june2019_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
airbnb_july2021_ppp_jit <- rjitter(airbnb_july2021_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
hotels_ppp_jit <- rjitter(hotels_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
tourism_ppp_jit <- rjitter(tourism_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

Check for duplicated points again.

```{r}
any(duplicated(airbnb_june2019_ppp_jit))
```

```{r}
any(duplicated(airbnb_july2021_ppp_jit))
```

```{r}
any(duplicated(hotels_ppp_jit))
```

```{r}
any(duplicated(tourism_ppp_jit))
```

There are no duplicated points after applying *jittering*.

### Creation of *owin* object

While analysing spatial point patterns, it is good to confine the analysis within a certain geographical area, which in this case will be Singapore boundary. ***owin*** in **spatstat** is specially designed to represent the polygonal region.

```{r}
sg_owin <- as(sg_sp, "owin")
```

Displaying the output object with the use of *plot()* function.

```{r}
plot(sg_owin)
```

Checking the summary of owin

```{r}
summary(sg_owin)
```

### Combining Point Events Object and owin object

Extracting the relevant events located within Singapore.

```{r}
mrtlrt_sg_ppp = mrtlrt_ppp[sg_owin]
airbnb_june2019_sg_ppp = airbnb_june2019_ppp_jit[sg_owin]
airbnb_july2021_sg_ppp = airbnb_july2021_ppp_jit[sg_owin]
hotels_sg_ppp = hotels_ppp_jit[sg_owin]
tourism_sg_ppp = tourism_ppp_jit[sg_owin]
```

Visualisation of ***ppp*** objects 

Using *plot()* function to visualise the ***ppp*** objects within Singapore's boundary

```{r fig.width=12, fig.height=8}
plot(mrtlrt_sg_ppp)
```

```{r fig.width=12, fig.height=8}
plot(airbnb_june2019_sg_ppp)
```

```{r fig.width=12, fig.height=8}
plot(airbnb_july2021_sg_ppp)
```

```{r fig.width=12, fig.height=8}
plot(hotels_sg_ppp)
```

```{r fig.width=12, fig.height=8}
plot(tourism_sg_ppp)
```

# Section A: Airbnb Distribution in 2019

## Exploratory Spatial Data Analysis

- Derive kernel density maps of Airbnb listings, hotels, MRT services, and tourist attractions.

- Using appropriate tmap functions, display the kernel density maps on openstreetmap of Singapore. Describe the spatial patterns revealed by the kernel density maps. Highlight the advantage of kernel density map over point map.

### Airbnb Listing June 2019

#### **Computing kernel density estimation using automatic bandwidth selection method**

Computing kernel density using *density()* of **spatstat** using *bw.diggle()*, a automatic bandwidth selection method. The default smoothing kernel is *gaussian*.

There are other bandwidth selection method such as *bw.ppl()* but I chosen to use *bw.diggle()* as it can identify single tight cluster patterns among random noise.

```{r}
kde_airbnb_june2019_sg_bw <- density(airbnb_june2019_sg_ppp,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

```{r}
plot(kde_airbnb_june2019_sg_bw)
```

After looking at the first graph, I realised the output value range is too small to comprehend. This is due to the default unit of measurement of svy21 in meter. Thus, it was computed in “number of points per square meter”.

#### **Rescalling KDE values to km(kilometer)**

```{r}
airbnb_june2019_sg_ppp_km <- rescale(airbnb_june2019_sg_ppp, 1000, 'km')
```

Now, the code can be re-run to get a better comprehension.

```{r}
kde_airbnb_june2019_sg_bw <- density(airbnb_june2019_sg_ppp_km,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

#### **Plotting of Kernel Density Estimation Maps**

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,2))
plot(kde_airbnb_june2019_sg_bw,  main = "KDE Map")
plot(airbnb_june2019_sg_ppp_km,  main = "Point Map")
```
Analysis: It can be seen that the Airbnb listings are mostly situated in the south-east region close to the central area. This can seen by the higher density in the map shown by the pink spots in the kernel density estimation map

Advantage of using kernel density map: Using the KDE map, the different shades of colour helps to visualise the density better as compared to point map. Point map only can only show the location of one place at the time, if too clustered together, it will just appear as a black patch and we would not be able to tell anything from it.

#### Converting KDE output into grid object.

Converting it so that it is suitable for mapping purposes.

```{r}
gridded_kde_airbnb_june2019_sg_bw <- as.SpatialGridDataFrame.im(kde_airbnb_june2019_sg_bw)
spplot(gridded_kde_airbnb_june2019_sg_bw)
```

The density map using grid shows a better visualisation of the listings as it is more visible to see those that are less clustered together as shown with the lighter shade of blue.

```{r}
kde_airbnb_june2019_sg_bw_m <- density(airbnb_june2019_sg_ppp,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

```{r}
gridded_kde_airbnb_june2019_sg_bw_m <- as.SpatialGridDataFrame.im(kde_airbnb_june2019_sg_bw_m)
```


#### **CONVERTING GRIDDED OUTPUT INTO RASTER**

Converting the gridded kernal density objects into RasterLayer object by using raster() of **raster** package.

```{r}
kde_airbnb_june2019_bw_raster <- raster(gridded_kde_airbnb_june2019_sg_bw_m)
kde_airbnb_june2019_bw_raster
```
The crs property is NA as shown above.

#### **ASSIGNING PROJECTION SYSTEMS** display the kernel density maps on openstreetmap of Singapore

Assigning the CRS information on kde_airbnb_june2019_bw_raster Raster Layer

```{r}
projection(kde_airbnb_june2019_bw_raster) <- CRS("+init=EPSG:3414")
kde_airbnb_june2019_bw_raster
```

The CRS property is now available.

Plotting the raster layer on openstreetmap using interactive mode

```{r}
tmap_mode("view")
tm_basemap("OpenStreetMap") +
tm_shape(kde_airbnb_june2019_bw_raster) + 
  tm_raster("v", palette = "Purples", alpha = 0.7)
```

Analysis: From the interactive map above, Airbnb can be seen in areas around Geylang which has a highest density of 0.00015 to 0.00020 units per square meter with the darkest shade of purple. Following that will be areas around Chinatown and Clarke Quay with a higher density of 0.00010 to 0.00015 units per square meter shown with the darker shade of purple.


### Hotels

#### **Rescalling KDE values to km(kilometer)**

```{r}
hotels_sg_ppp_km <- rescale(hotels_sg_ppp, 1000, 'km')
```

```{r}
kde_hotels_sg_bw <- density(hotels_sg_ppp_km,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

#### **Plotting of Kernel Density Estimation Maps**

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,2))
plot(kde_hotels_sg_bw,  main = "KDE Map")
plot(hotels_sg_ppp_km,  main = "Point Map")
```

#### Converting KDE output into grid object.

Converting it so that it is suitable for mapping purposes.

```{r}
gridded_kde_hotels_sg_bw <- as.SpatialGridDataFrame.im(kde_hotels_sg_bw)
spplot(gridded_kde_hotels_sg_bw)
```

The density map using grid shows a better visualisation of the listings as it is more visible to see those that are less clustered together as shown with the lighter shade of blue.

```{r}
kde_hotels_sg_bw_m <- density(hotels_sg_ppp,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

```{r}
gridded_kde_hotels_sg_bw_m <- as.SpatialGridDataFrame.im(kde_hotels_sg_bw_m)
```

#### **CONVERTING GRIDDED OUTPUT INTO RASTER**

Converting the gridded kernal density objects into RasterLayer object by using raster() of **raster** package.

```{r}
kde_hotels_bw_raster <- raster(gridded_kde_hotels_sg_bw_m)
kde_hotels_bw_raster
```
The crs property is NA as shown above.

#### **ASSIGNING PROJECTION SYSTEMS** display the kernel density maps on openstreetmap of Singapore

Assigning the CRS information on kde_airbnb_june2019_bw_raster Raster Layer

```{r}
projection(kde_hotels_bw_raster) <- CRS("+init=EPSG:3414")
kde_hotels_bw_raster
```

The CRS property is now available.

Plotting the raster layer on openstreetmap using interactive mode

```{r}
tmap_mode("view")
tm_basemap("OpenStreetMap") +
tm_shape(kde_hotels_bw_raster) + 
  tm_raster("v", palette = "Purples", alpha = 0.7)
```

Analysis: From the interactive map above, hotels can be seen that areas around Geylang and Clarke Quay has a highest density of 0.00015 to 0.00020 units per square meter. Following that will be areas around Lavender and Rocher with a higher density of hotel ranging from 0.00010 to 0.00015 units per square meter.


### MRT and LRT

#### **Rescalling KDE values to km(kilometer)**

```{r}
mrtlrt_sg_ppp_km <- rescale(mrtlrt_sg_ppp, 1000, 'km')
```

```{r}
kde_mrtlrt_sg_bw <- density(mrtlrt_sg_ppp_km,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

#### **Plotting of Kernel Density Estimation Maps**

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,2))
plot(kde_mrtlrt_sg_bw,  main = "KDE Map")
plot(mrtlrt_sg_ppp_km,  main = "Point Map")
```

#### Converting KDE output into grid object.

Converting it so that it is suitable for mapping purposes.

```{r}
gridded_kde_mrtlrt_sg_bw <- as.SpatialGridDataFrame.im(kde_mrtlrt_sg_bw)
spplot(gridded_kde_mrtlrt_sg_bw)
```

The density map using grid shows a better visualisation of the listings as it is more visible to see those that are less clustered together as shown with the lighter shade of blue.

```{r}
kde_mrtlrt_sg_bw_m <- density(mrtlrt_sg_ppp,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

```{r}
gridded_kde_mrtlrt_sg_bw_m <- as.SpatialGridDataFrame.im(kde_mrtlrt_sg_bw_m)
```

#### **CONVERTING GRIDDED OUTPUT INTO RASTER**

Converting the gridded kernal density objects into RasterLayer object by using raster() of **raster** package.

```{r}
kde_mrtlrt_bw_raster <- raster(gridded_kde_mrtlrt_sg_bw_m)
kde_mrtlrt_bw_raster
```
The crs property is NA as shown above.

#### **ASSIGNING PROJECTION SYSTEMS** display the kernel density maps on openstreetmap of Singapore

Assigning the CRS information on kde_airbnb_june2019_bw_raster Raster Layer

```{r}
projection(kde_mrtlrt_bw_raster) <- CRS("+init=EPSG:3414")
kde_mrtlrt_bw_raster
```

The CRS property is now available.

Plotting the raster layer on openstreetmap using interactive mode

```{r}
tmap_mode("view")
tm_basemap("OpenStreetMap") +
tm_shape(kde_mrtlrt_bw_raster) + 
  tm_raster("v", palette = "Purples", alpha = 0.7)  
```

Analysis: From the interactive map above, we can see that North-west, North-east and South-Central area has a higher density of MRT and LRT station, ranging 0.0000020 to 0.0000030.


### Tourist Attractions

#### **Rescalling KDE values to km(kilometer)**

```{r}
tourism_sg_ppp_km <- rescale(tourism_sg_ppp, 1000, 'km')
```

```{r}
kde_tourism_sg_bw <- density(tourism_sg_ppp_km,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

#### **Plotting of Kernel Density Estimation Maps**

```{r fig.width=12, fig.height=8}
par(mfrow=c(1,2))
plot(kde_tourism_sg_bw,  main = "KDE Map")
plot(tourism_sg_ppp_km,  main = "Point Map")
```

#### Converting KDE output into grid object.

Converting it so that it is suitable for mapping purposes.

```{r}
gridded_kde_tourism_sg_bw <- as.SpatialGridDataFrame.im(kde_tourism_sg_bw)
spplot(gridded_kde_tourism_sg_bw)
```

The density map using grid shows a better visualisation of the listings as it is more visible to see those that are less clustered together as shown with the lighter shade of blue.

```{r}
kde_tourism_sg_bw_m <- density(tourism_sg_ppp,
                                   sigma=bw.diggle,
                                   edge=TRUE,
                                   kernel="gaussian")
```

```{r}
gridded_kde_tourism_sg_bw_m <- as.SpatialGridDataFrame.im(kde_tourism_sg_bw_m)
```

#### **CONVERTING GRIDDED OUTPUT INTO RASTER**

Converting the gridded kernal density objects into RasterLayer object by using raster() of **raster** package.

```{r}
kde_tourism_bw_raster <- raster(gridded_kde_tourism_sg_bw_m)
kde_tourism_bw_raster
```
The crs property is NA as shown above.

#### **ASSIGNING PROJECTION SYSTEMS** display the kernel density maps on openstreetmap of Singapore

Assigning the CRS information on kde_airbnb_june2019_bw_raster Raster Layer

```{r}
projection(kde_tourism_bw_raster) <- CRS("+init=EPSG:3414")
kde_tourism_bw_raster
```

The CRS property is now available.

Plotting the raster layer on openstreetmap using interactive mode

```{r}
tmap_mode("view")
tm_basemap("OpenStreetMap") +
tm_shape(kde_tourism_bw_raster) + 
  tm_raster("v", palette = "Purples", alpha = 0.7)  
```

Analysis: There is high concentration of tourist attractions in the south-central, with a high density of 0.000015 to 0.000020.

### Comparing Spatial Point Patterns using KDE

The areas we will look into will be Anson, Bugis, Chinatown and Geylang as there are higher density of Airbnb

#### **EXTRACTING STUDY AREA**

```{r}
downtown = mpsz[mpsz@data$PLN_AREA_N == "DOWNTOWN CORE",] #for Anson
rochor = mpsz[mpsz@data$PLN_AREA_N == "ROCHOR",] # for Bugis
outram = mpsz[mpsz@data$PLN_AREA_N == "OUTRAM",] # for Chinatown
geylang = mpsz[mpsz@data$PLN_AREA_N == "GEYLANG",] # for Geylang

```

Plotting target planning areas

```{r}
par(mfrow=c(2,2))
plot(downtown, main = "Downtown")
plot(rochor, main = "Rochor")
plot(outram, main = "Outram")
plot(geylang, main = "Geylang")
```

#### **CONVERTING THE SPATIAL POINT DATA FRAME INTO GENERIC SP FORMAT**

Converting SpatialPolygonsDataFrame layers into generic spatialpolygons layers.

```{r}
downtown_sp = as(downtown, "SpatialPolygons")
rochor_sp = as(rochor, "SpatialPolygons")
outram_sp = as(outram, "SpatialPolygons")
geylang_sp = as(geylang, "SpatialPolygons")
```

#### **CREATING OWIN OBJECT**

Converting SpatialPolygons objects into owin objects that is required by **spatstat**

```{r}
downtown_owin = as(downtown_sp, "owin")
rochor_owin = as(rochor_sp, "owin")
outram_owin = as(outram_sp, "owin")
geylang_owin = as(geylang_sp, "owin")
```

#### **COMBINING AIRBNB JUNE 2019, HOTELS, MRTLRT, TOURISM POINTS AND THE STUDY AREA**

```{r}
airbnb_june2019_downtown_ppp = airbnb_june2019_ppp_jit[downtown_owin]
airbnb_june2019_rochor_ppp = airbnb_june2019_ppp_jit[rochor_owin]
airbnb_june2019_outram_ppp = airbnb_june2019_ppp_jit[outram_owin]
airbnb_june2019_geylang_ppp = airbnb_june2019_ppp_jit[geylang_owin]

hotels_downtown_ppp = hotels_ppp_jit[downtown_owin]
hotels_rochor_ppp = hotels_ppp_jit[rochor_owin]
hotels_outram_ppp = hotels_ppp_jit[outram_owin]
hotels_geylang_ppp = hotels_ppp_jit[geylang_owin]

mrtlrt_downtown_ppp = mrtlrt_ppp[downtown_owin]
mrtlrt_rochor_ppp = mrtlrt_ppp[rochor_owin]
mrtlrt_outram_ppp = mrtlrt_ppp[outram_owin]
mrtlrt_geylang_ppp = mrtlrt_ppp[geylang_owin]

tourism_downtown_ppp = tourism_ppp_jit[downtown_owin]
tourism_rochor_ppp = tourism_ppp_jit[rochor_owin]
tourism_outram_ppp = tourism_ppp_jit[outram_owin]
tourism_geylang_ppp = tourism_ppp_jit[geylang_owin]
```

Next, rescale() function is used to trasnform the unit of measurement from metre to kilometre

```{r}
airbnb_june2019_downtown_ppp_km = rescale(airbnb_june2019_downtown_ppp, 1000, "km")
airbnb_june2019_rochor_ppp_km = rescale(airbnb_june2019_rochor_ppp, 1000, "km")
airbnb_june2019_outram_ppp_km = rescale(airbnb_june2019_outram_ppp, 1000, "km")
airbnb_june2019_geylang_ppp_km = rescale(airbnb_june2019_geylang_ppp, 1000, "km")

hotels_downtown_ppp_km = rescale(hotels_downtown_ppp, 1000, "km")
hotels_rochor_ppp_km = rescale(hotels_rochor_ppp, 1000, "km")
hotels_outram_ppp_km = rescale(hotels_outram_ppp, 1000, "km")
hotels_geylang_ppp_km = rescale(hotels_geylang_ppp, 1000, "km")

mrtlrt_downtown_ppp_km = rescale(mrtlrt_downtown_ppp, 1000, "km")
mrtlrt_rochor_ppp_km = rescale(mrtlrt_rochor_ppp, 1000, "km")
mrtlrt_outram_ppp_km = rescale(mrtlrt_outram_ppp, 1000, "km")
mrtlrt_geylang_ppp_km = rescale(mrtlrt_geylang_ppp, 1000, "km")

tourism_downtown_ppp_km = rescale(tourism_downtown_ppp, 1000, "km")
tourism_rochor_ppp_km = rescale(tourism_rochor_ppp, 1000, "km")
tourism_outram_ppp_km = rescale(tourism_outram_ppp, 1000, "km")
tourism_geylang_ppp_km = rescale(tourism_geylang_ppp, 1000, "km")
```

#### **Airbnb June 2019** 

COMPUTING IDE

```{r fig.width=8, fig.height=6}
par(mfrow=c(2,2))
plot(density(airbnb_june2019_downtown_ppp_km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Downtown")
plot(density(airbnb_june2019_rochor_ppp_km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Rochor")
plot(density(airbnb_june2019_outram_ppp_km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Outram")
plot(density(airbnb_june2019_geylang_ppp, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Geylang")
```

#### **Are Airbnb listings are affected by location factors?**

**DOWNTOWN**

```{r fig.width=8, fig.height=6}
par(mfrow=c(2,2))
plot(airbnb_june2019_downtown_ppp_km, main = "Airbnb")
plot(hotels_downtown_ppp_km, main = "Hotel")
plot(mrtlrt_downtown_ppp_km, main = "MRT and LRT")
plot(tourism_downtown_ppp_km, main = "Tourist Attraction")
```

Analysis: Even though there are many hotels in the vicinity, there are more Airbnb as compared to hotels. It could be due to easy accessibility to many MRT stations and tourist attractions in the vicinity.

**ROCHOR**

```{r fig.width=8, fig.height=6}
par(mfrow=c(2,2))
plot(airbnb_june2019_rochor_ppp_km, main = "Airbnb")
plot(hotels_rochor_ppp_km, main = "Hotel")
plot(mrtlrt_rochor_ppp_km, main = "MRT and LRT")
plot(tourism_rochor_ppp_km, main = "Tourist Attraction")
```

Analysis: Even though there are many hotels in the Rochor, there are also many Airbnb as compared to hotels. It could be due to easy accessibility to many MRT stations nearby and tourist attractions in the vicinity.

**OUTRAM**

```{r fig.width=8, fig.height=6}
par(mfrow=c(2,2))
plot(airbnb_june2019_outram_ppp_km, main = "Airbnb")
plot(hotels_outram_ppp_km, main = "Hotel")
plot(mrtlrt_outram_ppp_km, main = "MRT and LRT")
plot(tourism_outram_ppp_km, main = "Tourist Attraction")
```

Analysis: Even though there are many hotels in the Rochor, there is more Airbnb as compared to hotels. However, there is very little MRT station and some tourist attractions in the vicinity.

**GEYLANG**

```{r fig.width=8, fig.height=6}
par(mfrow=c(2,2))
plot(airbnb_june2019_geylang_ppp_km, main = "Airbnb")
plot(hotels_geylang_ppp_km, main = "Hotel")
plot(mrtlrt_geylang_ppp_km, main = "MRT and LRT")
plot(tourism_geylang_ppp_km, main = "Tourist Attraction")
```

Analysis: There are many Airbnb in this area, with very little hotels nearby. It is quite accessible to MRT Stations but there is only one tourist attraction in the vicinity.

Overall: Even though the numbers of hotel in thee vicinity is high, there are still many Airbnb listing in all the areas. Every location was very accessible to MRT stations except Outram, but it didnt affect the number of Airbnb in the area. The same goes for tourist attractions, except for Geylang which only have one tourist attraction, but it did not affect the number of Airbnb in the area as well.

## Second-order Spatial Point Patterns Analysis

- Formulate the null hypothesis and alternative hypothesis and select the confidence level.

- Perform the test by using appropriate Second order spatial point patterns analysis technique.

- With reference to the analysis results, draw statistical conclusions.

### Nearest Neighbour Analysis

Perform the Clark-Evans test of aggregation for a spatial point pattern by using *clarkevans.test()* of **statspat**.

The test hypotheses are:

Ho = The distribution of Airbnb listings are randomly distributed.

H1= The distribution of Airbnb listings are not randomly distributed.

The 95% confident interval will be used.

### Testing spatial point patterns using Clark and Evans Test

```{r}
clarkevans.test(airbnb_june2019_sg_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

Analysis: The distribution of Airbnb in Singapore is found to be clustered as the R = 0.36 is < 1. The null hypothesis is rejected as the p-value = 0.01 is smaller than 0.05.

#### Clark and Evans Test: Downtown

```{r}
clarkevans.test(airbnb_june2019_downtown_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: The null hypothesis is rejected as the p-value = 0.002 is smaller than 0.05, showing that the Airbnb Listings in Downtown are not randomly distributed.

#### Clark and Evans Test: Rochor

```{r}
clarkevans.test(airbnb_june2019_rochor_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: The null hypothesis is rejected as the p-value = 0.002 is smaller than 0.05, showing that the Airbnb Listings in Rochor are not randomly distributed.

#### Clark and Evans Test: Outram

```{r}
clarkevans.test(airbnb_june2019_outram_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: The null hypothesis is rejected as the p-value = 0.002 is smaller than 0.05, showing that the Airbnb Listings in Outram are not randomly distributed.

#### Clark and Evans Test: Geylang

```{r}
clarkevans.test(airbnb_june2019_geylang_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Analysis: The null hypothesis is rejected as the p-value = 0.002 is smaller than 0.05, showing that the Airbnb Listings in Geylang are not randomly distributed.

Overall: ALl the areas shows that the Airbnb listing are not randomly distributed. AS Rochor and Outram has R value nearer to 1, we will focus on those 2 areas for further analysis.

